Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2023-11-20 lqg-strings.tex
Swapped y back for x in eq(25)
Added bimetric Polyakov EoMs to main file

Now, one might object and say that having an action which is the sum of Nambu-Goto actions for two different metrics seems strange. However, such actions have been extensively studied under the heading of ``bimetric gravity'', with the earliest work dating as far back as 1940 in two papers \cite{Rosen1940General,Rosen1940aGeneral} by Einstein's future collaborator Nathan Rosen (the `R' in `EPR' and `ER'). In 2010, bimetric gravity gained great popularity due to the seminal paper \cite{de-Rham2011Resummation} which showed that by introducing a second \emph{reference} metric into a model for massive gravity one can get rid of the Boulware-Deser ghost \cite{Boulware1972Can-Gravitation} which otherwise plagues theories with massive gravitons. The \emph{reference} metric introduced in the work \cite{de-Rham2011Resummation} was taken to be a flat metric. However later work \cite{Hassan2012Bimetric,Hassan2012Ghost-Free} showed that the second metric need not be flat and that an action of the following form:
\begin{align}\label{eqn:hassan-bimetric}
	S & = M_g^2 \int \diff^4 x \sqrt{-g} R^{(g)} + M_f^2 \int \diff^4 x \sqrt{-f} R^{(f)} \nonumber \\
	  & + 2 m^2 M^2_{eff} \int \diff^4 x \sqrt{-g}\sum_{n=0}^{4}\beta_n e_n (\sqrt{g^{-1} f}),
\end{align}
where $g_{\mu\nu}$ and $f_{\mu\nu}$ are \emph{arbitrary} metrics, $M_g^2, M_f^2$ are the Planck masses for the two sectors respectively. $R^{(g)}, R^{(f)}$ are the respective Ricci scalars for each sector. The third term is an interaction term which is responsible for making one of the graviton massive while the other remains massless. $M^2_{eff}$ is an \emph{effective} Planck mass given by:
\begin{equation}\label{eqn:hassan-effective-mpl}
	M^2_{eff} = \left( \frac{1}{M^2_g} + \frac{1}{M^2_f} \right)^{-1}
\end{equation}

@@

Using the static gauge $\tau = t$ and the transverse gauge with $\frac{\partial X}{\partial s} \cdot \frac{\partial X}{\partial t} = 0$, where $s$ is the length along the string parameter, we find that the new form for the conserved energy per infinitesimal piece of the string is the regular energy of the Nambu-Goto string times the correction factor
\begin{equation} \label{eqn:string-energy}
	E \diff \sigma = T \gamma _{v_{\perp}} \left( 1 + \frac{g \Delta}{2 (- h)} \right) \diff s ,
\end{equation}
and the form of the wave equation maintains it's shape as in
\begin{equation} \label{eqn:corrected-NG-wave-equation}
	\mu _{eff} \frac{\partial ^2 \Vec{X}}{\partial t^2} - \frac{\partial}{\partial s} \left[ T_{eff} \frac{\partial \Vec{X}}{\partial s} \right] = 0 ,
\end{equation}
where the effective mass density and effective tension are given by
\begin{align}
	\mu _{eff} & = \mu \gamma _{v_{\perp}} \left( 1 + \frac{g \Delta}{2 (- h)} \right) \label{eqn:effective-mass}            \\
	T_{eff}    & = \frac{T}{\gamma _{v_{\perp}}} \left( 1 + \frac{g \Delta}{2 (- h)} \right) . \label{eqn:effective-tension}
\end{align}

@@

Consider a string propagating in a target space whose geometry is described by two bulk metrics. There are two ways that a string with two metrics might arise. One possibility is that there are two sets of embedding fields $X^\mu$ and $Y^\mu$. As a result the induced metrics on the string worldsheet would also be different for both sets of embedding fields. However, in this case we would not be talking about \emph{one} string but two \emph{different} strings described by the respective embedding functions. One could then proceed along the lines of \eqref{eqn:hassan-bimetric} and introduce a third term for the worldsheet action which describes a coupling between the two sets of embedding fields, of the form:
\begin{equation}\label{eqn:ws-bimetric-interaction}
	S_{int} = - T_{eff} \int \diff^2 x\, \Phi(g,h) \Psi(\mb{X}, \mb{Y}).
\end{equation}
Here $\Phi(g,h)$ is some function of the \emph{induced} metrics $g_{ab}$ and $h_{ab}$ on the two worldsheets and $\Psi(\mb{X}, \mb{Y})$ is some function of the embedding fields. The resulting action would be of the form:
\begin{equation}\label{eqn:ws-bimetric-action-v1}
	S_{BNG} = -T \int \diff^2 y \, \sqrt{-h} - T' \int \diff^2 x \, \sqrt{-g} + S_{int}.
\end{equation}
Here the subscript $BNG$ stands for ``Bimetric Nambu-Goto''.

The second possibility is the following. Rather than considering two different sets of embedding fields, and consequently two different worldsheets, we can work with one set of embeddings $X^\mu(\tau,\sigma)$ and \emph{two different} bulk metrics $G_{\mu\nu}$ and $H_{\mu\nu}$. As a result we would now get two different induced metrics on the \emph{same} string worldsheet, given by:
\begin{subequations}\label{eqn:ws-induced-metrics}
	\begin{align}
		g_{ab} & = \frac{\partial X^\mu}{\partial x^a} \frac{\partial X^\nu}{\partial x^b} G_{\mu\nu}  \\
		h_{ab} & = \frac{\partial X^\mu}{\partial x^a} \frac{\partial X^\nu}{\partial x^b} H_{\mu\nu}.
	\end{align}
\end{subequations}
We can then write down the combined action for the string worldsheet as:
\begin{equation}\label{eqn:ws-bimetric-action-v2}
	S'_{BNG} = -T \int \diff^2 x\, \sqrt{-h} - T' \int \diff^2 x \, \sqrt{-g} + S'_{int},
\end{equation}
where now $S'_{int}[g,h,\mb{X}]$ is a term describing the interaction between the two metrics living on the \emph{same} worldsheet.

@@

The expression \eqref{eqn:ws-bimetric-action-v2} is not ideal because it does not retain any memory of the form of the bulk metrics, unless these are included in the interaction term $S'_{int}$. One way to manifestly exhibit the dependence of the respective worldsheet actions on the two bulk metrics is to work with the Polaykov action.

@@

@@

Let us begin by writing down the action of a string propagating in a background geometry which is specified not in terms of a metric, but in terms of a connection $A_\mu^{IJ}$ and vielbein fields $E_\mu^I$. Further we will require that the bulk connection and the bulk vielbein can be pulled back onto a connection $\omega _{a} ^{i j}$ and a dyad $e^a_i$, respectively, living on the string worldsheet $\Sigma$. In order for the bulk connection $A_\mu^{IJ}$, and its pullback $\omega _a ^{i j}$ on $\Sigma$, to have an effect on the string dynamics we will also promote the embedding fields $X^\mu$ to Lie algebra valued vector fields $X^{\mu} _{I}$. However, due to the fact that parallel transport of the embedding fields on the worldsheet should only involve the worldsheet connection $\omega _a ^{i j}$, we can replace $X^{\mu} _{I}$ with $X^{\mu} _{i}$, and to not lose the physical meaning of embedding fields as string coordinates, we include a new internal world-sheet field $v^{i}$, and define the string coordinates to be $\Tilde{X}^{\mu} \equiv X^{\mu} _{i} v^{i}$.
\begin{multline} \label{eqn:lqg-string-action}
	S_{PA} = - \frac{T}{2} \int \diff \tau \wedge \diff \sigma e e^{a} _{i} e^{b i} D_{a} X^{\mu} _{j} v^{j} \times \\
	\times D_{b} X^{\nu} _{k} v^{k} E_{\mu} ^{I} E_{\nu I} ,
\end{multline}
where $e = \det (e^{i} _{a})$ and $D_a$ is the covariant derivative with respect to the worldsheet connection, and is given by:
\begin{equation} \label{eqn:worldsheet-deriv}
	D_a X^{\mu} _{i} = \partial_a X^{\mu} _{i} - \omega_{a\,i}^{j} X^{\mu} _{j} ,
\end{equation}
where $\omega_{a\,j}^i$ is the pullback of the bulk connection to the worldsheet.

>>>>

Now, one might object and say that having an action which is the sum of Nambu-Goto actions for two different metrics seems strange. However, such actions have been extensively studied under the heading of ``bimetric gravity'', with the earliest work dating as far back as 1940 in two papers \cite{Rosen1940General,Rosen1940aGeneral} by Einstein's future collaborator Nathan Rosen (the `R' in `EPR' and `ER'). In 2010, bimetric gravity gained great popularity due to the seminal paper \cite{de-Rham2011Resummation} which showed that by introducing a second \emph{reference} metric into a model for massive gravity one can get rid of the Boulware-Deser ghost \cite{Boulware1972Can-Gravitation} which otherwise plagues theories with massive gravitons. The \emph{reference} metric introduced in the work \cite{de-Rham2011Resummation} was taken to be a flat metric. However later work \cite{Hassan2012Bimetric,Hassan2012Ghost-Free} showed that the second metric need not be flat and that an action of the following form:
\begin{align}\label{eqn:hassan-bimetric}
	S & = M_g^2 \int \diff^4 x \sqrt{-g} R^{(g)} + M_f^2 \int \diff^4 x \sqrt{-f} R^{(f)} \nonumber \\
	  & + 2 m^2 M^2_{eff} \int \diff^4 x \sqrt{-g}\sum_{n=0}^{4}\beta_n e_n (\sqrt{g^{-1} f}),
\end{align}
where $g_{\mu\nu}$ and $f_{\mu\nu}$ are \emph{arbitrary} metrics, $M_g^2, M_f^2$ are the Planck masses for the two sectors respectively. $R^{(g)}, R^{(f)}$ are the respective Ricci scalars for each sector. The third term is an interaction term which is responsible for making one of the gravitons massive while the other remains massless. $M^2_{eff}$ is an \emph{effective} Planck mass given by:
\begin{equation}\label{eqn:hassan-effective-mpl}
	M^2_{eff} = \left( \frac{1}{M^2_g} + \frac{1}{M^2_f} \right)^{-1}
\end{equation}

@@

Using the static gauge $\tau = t$ and the transverse gauge with $\frac{\partial X}{\partial s} \cdot \frac{\partial X}{\partial t} = 0$, where $s$ is the length along the string parameter, we find that the new form for the conserved energy per infinitesimal piece of the string is the regular energy of the Nambu-Goto string times the correction factor
\begin{equation} \label{eqn:string-energy}
	E \diff \sigma = T \gamma _{v_{\perp}} \left( 1 + \frac{g \Delta}{2 (- h)} \right) \diff s ,
\end{equation}
and the form of the wave equation maintains its shape as in
\begin{equation} \label{eqn:corrected-NG-wave-equation}
	\mu _{eff} \frac{\partial ^2 \Vec{X}}{\partial t^2} - \frac{\partial}{\partial s} \left[ T_{eff} \frac{\partial \Vec{X}}{\partial s} \right] = 0 ,
\end{equation}
where the effective mass density and effective tension are given by
\begin{align}
	\mu _{eff} & = \mu \gamma _{v_{\perp}} \left( 1 + \frac{g \Delta}{2 (- h)} \right) \label{eqn:effective-mass}            \\
	T_{eff}    & = \frac{T}{\gamma _{v_{\perp}}} \left( 1 + \frac{g \Delta}{2 (- h)} \right) . \label{eqn:effective-tension}
\end{align}

@@

Consider a string propagating in a target space whose geometry is described by two bulk metrics. There are two ways that a string with two metrics might arise. One possibility is that there are two sets of embedding fields $X^\mu$ and $Y^\mu$. As a result, the induced metrics on the string worldsheet would also be different for both sets of embedding fields. However, in this case, we would not be talking about \emph{one} string but two \emph{different} strings described by the respective embedding functions. One could then proceed along the lines of \eqref{eqn:hassan-bimetric} and introduce a third term for the worldsheet action which describes a coupling between the two sets of embedding fields, of the form:
\begin{equation}\label{eqn:ws-bimetric-interaction}
	S_{int} = - T_{eff} \int \diff^2 x\, \Phi(g,h) \Psi(\mb{X}, \mb{Y}).
\end{equation}
Here $\Phi(g,h)$ is some function of the \emph{induced} metrics $g_{ab}$ and $h_{ab}$ on the two worldsheets and $\Psi(\mb{X}, \mb{Y})$ is some function of the embedding fields. The resulting action would be of the form:
\begin{equation}\label{eqn:ws-bimetric-action-v1}
	S_{BNG} = -T \int \diff^2 x \, \sqrt{-h} - T' \int \diff^2 x \, \sqrt{-g} + S_{int}.
\end{equation}
Here the subscript $BNG$ stands for ``Bimetric Nambu-Goto''.

The second possibility is the following. Rather than considering two different sets of embedding fields, and consequently two different worldsheets, we can work with one set of embeddings $X^\mu(\tau,\sigma)$ and \emph{two different} bulk metrics $G_{\mu\nu}$ and $H_{\mu\nu}$. As a result, we would now get two different induced metrics on the \emph{same} string worldsheet, given by:
\begin{subequations}\label{eqn:ws-induced-metrics}
	\begin{align}
		g_{ab} & = \frac{\partial X^\mu}{\partial x^a} \frac{\partial X^\nu}{\partial x^b} G_{\mu\nu}  \\
		h_{ab} & = \frac{\partial X^\mu}{\partial x^a} \frac{\partial X^\nu}{\partial x^b} H_{\mu\nu}.
	\end{align}
\end{subequations}
We can then write down the combined action for the string worldsheet as:
\begin{equation}\label{eqn:ws-bimetric-action-v2}
	S'_{BNG} = -T \int \diff^2 x\, \sqrt{-h} - T' \int \diff^2 x \, \sqrt{-g} + S'_{int},
\end{equation}
where now $S'_{int}[g,h,\mb{X}]$ is a term describing the interaction between the two metrics living on the \emph{same} worldsheet.

@@

The expression \eqref{eqn:ws-bimetric-action-v2} is not ideal because it does not retain any memory of the form of the bulk metrics unless these are included in the interaction term $S'_{int}$. One way to manifestly exhibit the dependence of the respective worldsheet actions on the two bulk metrics is to work with the Polaykov action.

@@

\subsection{The Bimetric Polyakov Equations} \label{sec:bimetric-polyakov-eoms}

Following the usual procedure such as the one found in \cite{Tong2009Lectures}, we find two copies of the string energy-momentum tensor being set to zero, which can be used to find an expression for the induced metrics:
\begin{equation} \label{eqn:variation-wrt-g}
	\frac{\delta S_{P B}}{\delta g^{c d}} \overset{!}{=} 0
\end{equation}
\begin{equation*}
	\downarrow
\end{equation*}
\begin{multline} \label{eqn:g-em-tensor}
	T^{(G)} _{c d} := G_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\nu} - \\
	- \frac{1}{2} g_{c d} g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = 0
\end{multline}
\begin{equation*}
	\Downarrow
\end{equation*}
\begin{equation} \label{eqn:g-expression}
	g_{c d} = 2 f^{(G)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} G_{\mu \nu} ,
\end{equation}
where the function $f^{(G)}$ is given by
\begin{equation} \label{eqn:g's-f-function}
	\frac{1}{f^{(G)}} = g^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} G_{\mu \nu},
\end{equation}
and completely analogously for the $H_{\mu \nu}$ metric,
\begin{equation} \label{variation-wrt-h}
	\frac{\delta S_{P B}}{\delta h^{c d}} \overset{!}{=} 0
\end{equation}
\begin{equation*}
	\downarrow
\end{equation*}
\begin{multline} \label{h-em-tensor}
	T^{(H)} _{c d} := H_{\mu \nu} \bigg( \partial _{c} X^{\mu} \partial _{d} X^{\mu} - \\
	- \frac{1}{2} h_{c d} h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \bigg) = 0
\end{multline}
\begin{equation*}
	\Downarrow
\end{equation*}
\begin{equation} \label{h-expression}
	h_{c d} = 2 f^{(H)} \partial _{c} X^{\mu} \partial _{d} X^{\nu} H_{\mu \nu} ,
\end{equation}
with
\begin{equation} \label{h's-f-function}
	\frac{1}{f^{(H)}} = h^{a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} H_{\mu \nu} .
\end{equation}

As for the embbeding fields equation, we get a wave equation for the combined metric
\begin{equation} \label{eqn:variation-wrt-X}
	\frac{\delta S_{P B}}{\delta X^{\lambda}} \overset{!}{=} 0
\end{equation}
\begin{equation*}
	\downarrow
\end{equation*}
\begin{equation} \label{eqn:bimetric-wave-equation}
	f^{\prime a b} \partial _{a} \partial _{b} X^{\mu} F_{\mu \lambda} = \frac{1}{2} f^{\prime a b} \partial _{a} X^{\mu} \partial _{b} X^{\nu} \frac{\partial F_{\mu \nu}}{\partial X^{\lambda}} ,
\end{equation}
where
\begin{equation} \label{eqn:coupled-ws-metric}
	f^{\prime a b} & = T \sqrt{- g} g^{a b} + T' \sqrt{- h} h^{a b}
\end{equation}
\begin{equation} \label{eqn:combined-bulk-metric}
	F_{\mu \nu}    & = G_{\mu \nu} + H_{\mu \nu} ,
\end{equation}
in which we add a prime in the coupled metric $f^{\prime a b}$ since it is not quite the metric induced by the combined bulk metric $F_{\mu \nu}$.

@@

Let us begin by writing down the action of a string propagating in a background geometry which is specified not in terms of a metric, but in terms of a connection $A_\mu^{IJ}$ and vielbein fields $E_\mu^I$. Further, we will require that the bulk connection and the bulk vielbein can be pulled back onto a connection $\omega _{a} ^{i j}$ and a dyad $e^a_i$, respectively, living on the string worldsheet $\Sigma$. In order for the bulk connection $A_\mu^{IJ}$, and its pullback $\omega _a ^{i j}$ on $\Sigma$, to have an effect on the string dynamics we will also promote the embedding fields $X^\mu$ to Lie algebra valued vector fields $X^{\mu} _{I}$. However, due to the fact that parallel transport of the embedding fields on the worldsheet should only involve the worldsheet connection $\omega _a ^{i j}$, we can replace $X^{\mu} _{I}$ with $X^{\mu} _{i}$, and to not lose the physical meaning of embedding fields as string coordinates, we include a new internal world-sheet field $v^{i}$, and define the string coordinates to be $\Tilde{X}^{\mu} \equiv X^{\mu} _{i} v^{i}$.
\begin{multline} \label{eqn:lqg-string-action}
	S_{PA} = - \frac{T}{2} \int \diff \tau \wedge \diff \sigma e e^{a} _{i} e^{b i} D_{a} X^{\mu} _{j} v^{j} \times \\
	\times D_{b} X^{\nu} _{k} v^{k} E_{\mu} ^{I} E_{\nu I} ,
\end{multline}
where $e = \det (e^{i} _{a})$ and $D_a$ is the covariant derivative with respect to the worldsheet connection, and is given by:
\begin{equation} \label{eqn:worldsheet-deriv}
	D_a X^{\mu} _{i} = \partial_a X^{\mu} _{i} - \omega_{a\,i}^{j} X^{\mu} _{j} ,
\end{equation}
where $\omega_{a\,j}^i$ is the pullback of the bulk connection to the worldsheet.
